timesliced_random_cell_selection

### 2018-12-19 src/timesliced_random_cell_selection.Rmd
random cell generation (10k)

```{r}
bckrnd_folder <- "../output/bio/background/raw/"      # path to folder that holds multiple .csv files
bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
# read in each .csv file in file_list and create a data frame with the same name as the .csv file
for (i in 1:length(bckrnd_list)){
  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
  sel <- sample_n(sel, 20000) #where 20000 = number of rows to sample (more than 10k as some points will have NA values due to depth - I will select 10000 of these without NA values to create a large sample as per maxent)
  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(sel)
```

@@

```{r}
rand_folder <- "../output/bio/background/rand/"      # path to folder that holds multiple .csv files
rand_list <- list.files(path = rand_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
for(i in 1:length(rand_list)){
  randtemp <- read.csv(paste(rand_folder, rand_list[i], sep=''))
  randtemp$year <- substr(rand_list[i], 1, 4) #the year is characters 1-4 in the filename
  randtemp$month <- substr(rand_list[i], 6, 7) #the month is characters 6 & 7 in the filename (except where the month is a single digit)
  write.csv(randtemp, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(randtemp)
```
note that in the output temp file in R shows a . after the month if that month is a single digit. The .csv outputs don't

@@

First you need to put the background (background_all) into a spatialpointsdataframe.
```{r}
xy <- background_all[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
head(background_all_sp)
```

@@

    }
}
write.csv(background_all_sp, "../output/bio/background_complete.csv", row.names = FALSE)
background_all <- as.data.frame(background_all_sp)
head(background_all)
```

@@

>>>>

```{r}
bckrnd_folder <- "../output/bio/background/raw/"      # path to folder that holds multiple .csv files
bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand20k/" #where the files are to be saved to
# read in each .csv file in file_list and create a data frame with the same name as the .csv file
for (i in 1:length(bckrnd_list)){
  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
  sel <- sample_n(sel, 20000) #where 20000 = number of rows to sample (more than 10k as some points will have NA values due to depth - I will select 10000 of these without NA values to create a large sample as per maxent)
  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(sel)
```

@@

```{r}
rand_folder <- "../output/bio/background/rand20k/"      # path to folder that holds multiple .csv files
rand_list <- list.files(path = rand_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand20k/" #where the files are to be saved to
for(i in 1:length(rand_list)){
  randtemp <- read.csv(paste(rand_folder, rand_list[i], sep=''))
  randtemp$year <- substr(rand_list[i], 1, 4) #the year is characters 1-4 in the filename
  randtemp$month <- substr(rand_list[i], 6, 7) #the month is characters 6 & 7 in the filename (except where the month is a single digit)
  write.csv(randtemp, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(randtemp)
```
note that in the output temp file in R shows a . after the month if that month is a single digit. The .csv outputs don't

@@

First you need to create columns for the data to be added to (otherwise the loop doesn't always work - notably on one PC but not another)

```{r}
background_all$temp_depth <- NA
background_all$temp_surface <- NA
background_all$salinity_depth <- NA
background_all$salinity_surface <- NA
background_all$chl_depth <- NA
background_all$chl_surface <- NA
background_all$o2_depth <- NA
background_all$o2_surface <- NA
background_all$mlp_surface <- NA
background_all$ssh_surface <- NA
```

Second you need to put the background (background_all) into a spatialpointsdataframe.
```{r}
xy <- background_all[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
head(background_all_sp)
```

@@

    }
}
write.csv(background_all_sp, "../output/bio/background_complete.csv", row.names = FALSE)
background_all <- as.data.frame(background_all_sp)
head(background_all_sp)
```

@@

# remove rows with NA values at depth

first rbind... (this you can delete if you had run the code at once)
```{r}
back1998 <- read.csv("../output/bio/background_complete_1998.csv", header = TRUE)
back1999 <- read.csv("../output/bio/background_complete_1999.csv", header = TRUE)
back2000_2003 <- read.csv("../output/bio/background_complete_2000_2003.csv", header = TRUE)
back2004_2006 <- read.csv("../output/bio/background_complete_2004_2006.csv", header = TRUE)
back2007_2011 <- read.csv("../output/bio/background_complete_2007_2011.csv", header = TRUE)
back2012_2016 <- read.csv("../output/bio/background_complete_2012_2016.csv", header = TRUE)

background_all <- rbind(back1998, back1999, back2000_2003, back2004_2006, back2007_2011, back2012_2016)
head(background_all)
```

remove all na values

```{r}
backna <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
backna
background_all[complete.cases(background_all), ] #removes NA values
write.csv(background_all, "../output/bio/background_complete_na.csv", row.names = FALSE)
```

randomly select 10000 rows from each timeslice
```{r}
#background_all <- read.csv("../output/bio/background_complete_na.csv", header = TRUE)
#split_cell_back <- split(background_all, list(background_all$year, background_all$month), drop = TRUE)

no_timeslices <- length(split_cell_back) # how many levels (time slices) are in the list
bckoutput <- "../output/bio/background/rand20k/no_na/" #where the files are to be saved to
for (i in 1:no_timeslices){
  timeslice <- split_cell_back[[i]]
  timeslice <- as.data.frame(timeslice)
  write.csv(timeslice, paste0(bckoutput, names(lapply(split_cell_back, names))[i], ".csv"), row.names = FALSE) 
}
```

```{r}
bckrnd_folder <- "../output/bio/background/rand20k/no_na/"      # path to folder that holds multiple .csv files
bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- ". ./output/bio/background/rand20k/no_na_10krand/" #where the files are to be saved to
# read in each .csv file in file_list and create a data frame with the same name as the .csv file
for (i in 1:length(bckrnd_list)){
  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
  sel <- sample_n(sel, 10000) #where 10000 = number of rows to sample 
  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(sel)
```
